Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-03-20 paper-1/elife.tex
Better fig name: 9 -> space

\begin{figure}
    \includegraphics[width=\linewidth]{fig9}
    \caption{%
        Suppression acts in spatially distinguishable subpopulations.
        \textbf{A} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial). Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin. Left, spatial illustration, using all data with a fictive stimulus location. Right, same data plotted as lines. The vertical line indicates the chosen local/global divider.
        \textbf{B} As in panel A, but showing $\T{A}{}$ without normalization.
        \textbf{C} As in panel A, but showing $\D{A}{}$ without normalization.
        \textbf{D} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global). Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot. Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
        \textbf{E} Corresponding mean $\D{A}{}$ in local and global portions of the network.
    }
    \label{fig:9}
\end{figure}

Finally, we sought to confirm whether ``local'' and ``global'' suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network. To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance. \FIG{9}A-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}. Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site. For thresholds, $\T[std]{A}{}$ (\FIG{9}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms. A similar pattern emerged with $\D{A}{}$ (\FIG{9}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.

To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000. We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}D. In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression. Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both $\T{A}{}$ (T = 1.09e+03, p = 1.37e-11) and $\D{A}{}$ (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence. Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for $\T{A}{}$ (T = 831, p = 1.15e-13) and $\D{A}{}$ (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.

>>>>

\begin{figure}
    \includegraphics[width=\linewidth]{fig-space}
    \caption{%
        Suppression acts in spatially distinguishable subpopulations.
        \textbf{A} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial). Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin. Left, spatial illustration, using all data with a fictive stimulus location. Right, same data plotted as lines. The vertical line indicates the chosen local/global divider.
        \textbf{B} As in panel A, but showing $\T{A}{}$ without normalization.
        \textbf{C} As in panel A, but showing $\D{A}{}$ without normalization.
        \textbf{D} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global). Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot. Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
        \textbf{E} Corresponding mean $\D{A}{}$ in local and global portions of the network.
    }
    \label{fig:space}
\end{figure}

Finally, we sought to confirm whether ``local'' and ``global'' suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network. To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance. \FIG{space}A-C shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}. Responses, normalized by the response to stimulation in the recovered networks (\FIG{space}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site. For thresholds, $\T[std]{A}{}$ (\FIG{space}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms. A similar pattern emerged with $\D{A}{}$ (\FIG{space}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.

To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000. We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{space}D. In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression. Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both $\T{A}{}$ (T = 1.09e+03, p = 1.37e-11) and $\D{A}{}$ (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence. Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for $\T{A}{}$ (T = 831, p = 1.15e-13) and $\D{A}{}$ (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 9 -> space

\begin{figure*}%
    \centering
    \includegraphics{fig9.pdf}
    \caption{}
    \label{fig:9}
\end{figure*}
\textbf{Fig. 9} Suppression acts in spatially distinguishable subpopulations.
\textbf{a} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial). Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin. Left, spatial illustration, using all data with a fictive stimulus location. Right, same data plotted as lines. The vertical line indicates the chosen local/global divider.
\textbf{b} As in panel a, but showing $\T{A}{}$ without normalization.
\textbf{c} As in panel a, but showing $\D{A}{}$ without normalization.
\textbf{d} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global). Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot. Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
\textbf{e} Corresponding mean $\D{A}{}$ in local and global portions of the network.

Finally, we sought to confirm whether ``local'' and ``global'' suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network. To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance. \FIG{9}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}. Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site. For thresholds, $\T[std]{A}{}$ (\FIG{9}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms. A similar pattern emerged with $\D{A}{}$ (\FIG{9}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.

To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000. We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}d. In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression. Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both $\T{A}{}$ (T = 1.09e+03, p = 1.37e-11) and $\D{A}{}$ (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence. Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for $\T{A}{}$ (T = 831, p = 1.15e-13) and $\D{A}{}$ (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.

>>>>

\begin{figure*}%
    \centering
    \includegraphics{fig-space.pdf}
    \caption{}
    \label{fig:space}
\end{figure*}
\textbf{Fig. 9} Suppression acts in spatially distinguishable subpopulations.
\textbf{a} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial). Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin. Left, spatial illustration, using all data with a fictive stimulus location. Right, same data plotted as lines. The vertical line indicates the chosen local/global divider.
\textbf{b} As in panel a, but showing $\T{A}{}$ without normalization.
\textbf{c} As in panel a, but showing $\D{A}{}$ without normalization.
\textbf{d} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global). Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot. Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
\textbf{e} Corresponding mean $\D{A}{}$ in local and global portions of the network.

Finally, we sought to confirm whether ``local'' and ``global'' suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network. To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance. \FIG{space}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}. Responses, normalized by the response to stimulation in the recovered networks (\FIG{space}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site. For thresholds, $\T[std]{A}{}$ (\FIG{space}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms. A similar pattern emerged with $\D{A}{}$ (\FIG{space}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.

To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000. We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{space}d. In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression. Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both $\T{A}{}$ (T = 1.09e+03, p = 1.37e-11) and $\D{A}{}$ (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence. Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for $\T{A}{}$ (T = 831, p = 1.15e-13) and $\D{A}{}$ (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.
